---
title: "Linear Regression 1"
author: "John Yuill"
date: "October 23, 2016"
output: html_document
---

### Purpose
* collect practices around correlation and regression
* primary initial source is Marin Stats R video on YT
    + https://www.youtube.com/watch?v=66z_MRwtFJM

#### Collect data and have a look
```{r}
library(ggplot2)


lungcap <- read.csv("data-source/LungCapData.csv",header=TRUE, stringsAsFactors=FALSE)
str(lungcap)

plot(lungcap$Age, lungcap$LungCap, main="Scatterplot")
ggplot(data=lungcap, aes(x=Age,y=LungCap))+geom_point()+theme_bw()

```

#### Check Correlation and Create Linear Model

* Check correlation
```{r}
cor(lungcap$Age,lungcap$LungCap)
```

* Generate a linear model
```{r}
mod <- lm(lungcap$LungCap ~ lungcap$Age)
summary(mod)
```

* [sidebar] determine attributes available for the mod object
```{r}
attributes(mod)
```

* check coefficients 
```{r}
mod$coef
coef(mod)
```

* get confidence intervals
    + 95% is default
```{r}
confint(mod)
```
    + adjust to 99%
```{r}
confint(mod, level=0.99)
```

* scatterplot with linear regression line
```{r}
plot(lungcap$Age, lungcap$LungCap, main="Scatterplot")
abline(mod, col=2, lwd=3)

ggplot(data=lungcap, aes(x=Age,y=LungCap))+geom_point()+
theme_bw()+
geom_smooth(method=lm, se=TRUE)

```

#### Check validity of assumptions

Assumptions
1. Y-values (or the errors) are independent
    + requires knowledge of the study design or data collection
2. Y-values can be expressed as a linear function of the X variable
3. Variation of observations around the regression line (residual standard errors (SE)) is constant (homoscedasticity)
4. For given value of X, Y values (or the error) are normally distributed

* validation of #1 requires knowledge of the study design or data collected
* others can be checked by examining the residuals or errors

##### R has built-in regression diagnostic plots

Diagnostic plots  

* highly useful for multiple regression, where the model can't be captured in a scatter plot

* unless mfrow is set, plots require hitting 'return' in console  
* won't run in R Markdown - will cause error
```{r}
par(mfrow=c(2,2))
plot(mod)


```

The Plots:  

1. Residuals vs Fitted
    + should see no real pattern
    + line should be flat - indicates linearity
    + heteroskedasticity will have a 'megaphone' shape, wider to the right
2. Normal Q-Q
    + if errors are normally distributed, points should fall along line
3. Scale-Location 
    + line should be flat - if not, indicates issues with linearity
4. Residuals vs Leverage
    + 


```{r}

```

